Layered Formal Verification of a TCP/IP Stack

### 2021-06-02 SecDev21/paper.tex
fixing e.g.
reviewing the conclusion

In total, for this work, 50 subprograms have been translated from C to SPARK. This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g\.~ contracts) used to represent and prove the protocol's specification. The overall proving time with GNATprove is 9m and 54s on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.

@@

We selectively applied formal verification to a layer of a TCP/IP library, by
translating TCP user functions into SPARK, and modelling asynchronous events
with ghost code. This allowed us to detect two bugs in the original C
implementation, related to memory management and concurrency. This is a step
towards a more secure implementation. The translated code has been proven to be
free of run-time errors and to respect the transitions of the TCP state machine
as specified in RFC 793.

These results depend upon other layers in the C programming language, which
were not formally verified to the same level. Instead, we used symbolic
execution with KLEE to formally verify the contracts used in SPARK to model the
corresponding behaviors. The principal function of those layers
is to format packets before they are sent, or parse incoming packets, check their
integrity and transmit the resulting payload to the corresponding upper-level
layer. This processing part can be a source of errors and bugs as explain
in~\cite{Reiher2019RecordFluxFM}.
As an alternative to manual translation into SPARK, we envision 
using the RecordFlux DSL to specify the parser and printer of the packets of the
different protocols, and to generate provable SPARK code from this specification.
This would complete our current verification of the CycloneTCP stack towards
the lower layers.

>>>>

In total, for this work, 50 subprograms have been translated from C to SPARK. This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove the protocol's specification. The overall proving time with GNATprove is 9m and 54s on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.

@@

We selectively applied formal verification to a TCP/IP library layer by translating TCP user functions into SPARK and modeling asynchronous events with ghost code. This enabled the detection of two bugs in the original C implementation related to memory management and concurrency. The work is a step towards a more secure implementation. The translated code has been proven to be free of run-time errors and to respect the transitions of the TCP state machine, specified in RFC 793 \cite{rfc793}.

These results depend upon other layers in the C programming language, which were not formally verified to the same level. Instead, we used symbolic execution with KLEE to formally verify the contracts used in SPARK to model the corresponding behaviors. The principal function of those layers is to format packets before they are sent, or parse incoming packets, check their integrity, and transmit the resulting payload to the corresponding upper-level layer. This processing part can be a source of errors and bugs as explain in~\cite{Reiher2019RecordFluxFM}. As an alternative to manual translation into SPARK, we envision using the RecordFlux DSL \cite{Reiher2019RecordFluxFM} to specify the parser and printer of the different protocols' packets and generate provable SPARK code from this specification. This would complete our current verification of the CycloneTCP stack towards the lower layers.
